---
title: "Advanced ridigbio"
author: "Matthew Collins"
date: "October 5, 2015"
output: html_document
---

> ## Learning Objectives
>
> * Filter and select data using ridigbio
> * Read API documentation
> * Broaden understanding of API and package posabilities

## Using ridigbio and the API to do it

The CSV file we just wrote probably only had one line in it when we finished 
removing all the data we could not use. But, we had to download 10 records over
the internet to get enough data to find one good record. What if we wanted 100 
good records? How many would we need to download? Can we know?

A better approach than filtering data yourself is to see if the API or R 
package will do it for you. In the case of the iDigBio API and the R package,
the answer is "yes!" And the filtering happens on the server at iDigBio so only
good records are transfered to you.

### Reading the Documentation

In order to know how to use ridigbio to filter out the records we don't want,
we need to read the documentation for the package. In R, you can put a `?`
(question mark) in front of a package or function name on the console line.
In the lower right section, the Help tab will open to help for the function.

```{r, eval=FALSE, purl=FALSE}
?ridigbio
?idig_search_records
```

The most useful parameter to the function is `rq` which is "iDigBio query in
nested list format". Later in the help page there is a link to the specification
for this format:

[https://github.com/iDigBio/idigbio-search-api/wiki/Query-Format](https://github.com/iDigBio/idigbio-search-api/wiki/Query-Format)

You may recognize this query format as a JSON object. Recall that the iDigBio
API takes JSON and returns JSON.

### Constructing a JSON object in R

```{r, eval=FALSE, purl=FALSE}
query <- list(scientificname=list(type="exists"), 
              geopoint=list(type="exists"),
              genus="acer")
```

### Requiring a Field to be Populated

```{r, eval=FALSE, purl=FALSE}
df <- idig_search_records(limit=10, rq=query)
```

### Limiting the Columns Returned

```{r, eval=FALSE, purl=FALSE}
field_list <- c("scientificname", "geopoint")
df <- idig_search_records(limit=10, rq=query, fields=field_list)
```

## Bonus: Everyone has an API and a package

Data access APIs and packages in R are common. If you want some kind of data,
try seeing if the data provider has an API or package for your favorite
programming language.

Some ideas:

1. [rOpenSci](https://ropensci.org/packages/)
1. [Machine learning on CRAN](https://cran.r-project.org/web/views/MachineLearning.html)
1. [Bioconductor distribution for bioinformatics](https://www.bioconductor.org/)